---
title: "Chapter 2 module prep"
author: "Mohammed Ali"
date: "2023-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dbdataset)
library(glue)
library(canvasXpress)
library(dplyr)
library(lubridate)
library(tidyr)

drugs <- readr::type_convert(drugbank$drugs$general_information, guess_integer = T)

statistical_mode <- function(x) {
  unique_x    <- unique(x)
  frequencies <- tabulate(match(x, unique_x))
  unique_x[frequencies == max(frequencies)]
}

get_variable_type <- function(target_variable, categories_number = 5) {
    var_type <- "character"
    
    if (NROW(target_variable) == length(unique(target_variable))) {
        var_type <- "ID"
    } else if (length(unique(target_variable)) <= categories_number) {
        var_type <- "factor"
    } else if (is.numeric(target_variable)) {
        #var_type <- "numeric" #EX: make it more specific
        var_type <- "double"
        if (all(drugs$average_mass == round(drugs$average_mass))) {
            var_type <- "integer"
        }
    } else if (inherits(target_variable, c("Date", "POSIXt"))) {
        var_type <- "Date"
    }
    
    var_type
}
```

## Dataset Overview

- Number of observations: `r NROW(drugs)`
- Number of variables	: `r NCOL(drugs)`

- Variables types:
```{r drugs}
 # sapply(drugs, get_variable_type)
```
 EX: control what to display

## Numerical

- Variable type: `r get_variable_type(drugs$average_mass)`
- Number of missing obs.: `r glue("{round(sum(is.na(drugs$average_mass))/NROW(drugs$average_mass), 4) * 100}%")`
- Number of unique values: `r length(unique(na.omit(drugs$average_mass)))`
- Median: `r round(median(drugs$average_mass, na.rm = TRUE), 2)`
- Mode: `r round(statistical_mode(na.omit(drugs$average_mass)), 2)`
- Variance: `r round(var(drugs$average_mass, na.rm = TRUE), 2)`
- SD: `r round(sd(drugs$average_mass, na.rm = TRUE), 2)`
- 1st quantile: `r quantile(drugs$average_mass, 1/4, na.rm = TRUE)`
- 3rd quantile: `r quantile(drugs$average_mass, 3/4, na.rm = TRUE)`
- Max: `r max(drugs$average_mass, na.rm = TRUE)`
- Min: `r min(drugs$average_mass, na.rm = TRUE)`

```{r hist}
d <- data.frame("Average Mass" = drugs$average_mass)
canvasXpress(
    data=d,
    graphType="Scatter2D",
    histogramBins=10,
    title="Average Mass",
    xAxisTitle="Average Mass",
    yAxisTitle="Count",
    showLegend = FALSE,
    histogramData=TRUE
)

canvasXpress(
    data=d,
    graphType="Scatter2D",
    histogramBins=10,
    title="Average Mass",
    xAxisTitle="Average Mass",
    yAxisTitle="Count",
    showLegend = FALSE,
    hideHistogram=TRUE,
    histogramData=TRUE,
    showHistogramDensity=TRUE
)
```

## Date
- Variable type: `r get_variable_type(drugs$created)`
- Number of missing obs.: `r glue("{round(sum(is.na(drugs$created))/NROW(drugs$created), 4) * 100}%")`
- Number of unique values: `r length(unique(na.omit(drugs$created)))`
- Median: `r round(median(drugs$created, na.rm = TRUE), 2)`
- Mode: `r round(statistical_mode(na.omit(drugs$created)), 2)``
- 1st quantile: `r quantile(drugs$created, 1/4, na.rm = TRUE, type = 1)`
- 3rd quantile: `r quantile(drugs$created, 3/4, na.rm = TRUE, type = 1)`
- Max: `r max(drugs$created, na.rm = TRUE)`
- Min: `r min(drugs$created, na.rm = TRUE)`

```{r date_bar}
# group by whole date
create_date <- drugs %>% 
    select(primary_key, created) %>%
    group_by(created) %>% 
    summarise(no_drugs = n()) %>% as.data.frame()
rownames(create_date) <- create_date$created
create_date$created   <- NULL
canvasXpress(
    data=create_date,
    graphOrientation="vertical",
    graphType="Bar",
    showLegend=FALSE,
    smpLabelRotate=90,
    smpTitle="Samples",
    title="Bar Graph Title",
    xAxisTitle="Value"
)
# group by whole year
create_date <- drugs %>%
    mutate(created = year(created)) %>% 
    select(primary_key, created) %>%
    group_by(created) %>% 
    summarise(no_drugs = n()) %>% as.data.frame()
rownames(create_date) <- create_date$created
create_date$created   <- NULL
canvasXpress(
    data=create_date,
    graphOrientation="vertical",
    graphType="Bar",
    showLegend=FALSE,
    smpLabelRotate=90,
    smpTitle="Samples",
    title="Bar Graph Title",
    xAxisTitle="Value"
)

# group by month
create_date <- drugs %>%
    mutate(created = month(created)) %>% 
    select(primary_key, created) %>%
    group_by(created) %>% 
    summarise(no_drugs = n()) %>% as.data.frame()
rownames(create_date) <- create_date$created
create_date$created   <- NULL
canvasXpress(
    data=create_date,
    graphOrientation="vertical",
    graphType="Bar",
    showLegend=FALSE,
    smpLabelRotate=90,
    smpTitle="Samples",
    title="Bar Graph Title",
    xAxisTitle="Value"
)
```


## Factor
- Variable type: `r get_variable_type(drugs$state)`
- Number of missing obs.: `r glue("{round(sum(is.na(drugs$state))/NROW(drugs$state), 4) * 100}%")`
- Number of unique values: `r length(unique(na.omit(drugs$state)))`
- Mode: `r statistical_mode(na.omit(drugs$state))`

```{r bar}
create_date <- drugs %>% 
    select(primary_key, state) %>%
    drop_na(state) %>% 
    group_by(state) %>% 
    summarise(no_drugs = n()) %>% as.data.frame()
rownames(create_date) <- create_date$state
create_date$state   <- NULL
canvasXpress(
    data=create_date,
    graphOrientation="vertical",
    graphType="Bar",
    showLegend=FALSE,
    smpLabelRotate=90,
    smpTitle="Samples",
    title="Bar Graph Title",
    xAxisTitle="Value"
)
```

## Character
## ID


